Collection of BAAD plots

### 2014-06-19 Rmd/baadplotscollection.Rmd
Added more figures, also Daniel's plots of all data. html compiles fine, many small details fixed.

source("load.R")

@@

pipe_bh <- summarize(group_by(dat, dataset, species), 
           alf_stbh=mean(a.lf / a.stbh),
           mlf_stbh=mean(m.lf / a.stbh),
           sla = mean(a.lf / m.lf),
           pft=first(pft),
           MAP=mean(MAP),
           MAT=mean(MAT),
           growingCondition=first(growingCondition),
           vegetation=first(vegetation))

# Basal pipemodel relationships
datl <- studyWithVars(baad, c("a.stba", "a.lf","m.lf"), returnwhat="list")

dat$pft <- as.factor(dat$pft)
pipe_bh$pft <- as.factor(pipe_bh$pft)

pipe_ba <- summarize(group_by(dat, dataset, species), 
                     alf_stba=mean(a.lf / a.stba),
                     mlf_stba=mean(m.lf / a.stba),
                     pft=first(pft),
                     MAP=mean(MAP),
                     MAT=mean(MAT),
                     growingCondition=first(growingCondition),
                     vegetation=first(vegetation))

@@

  N <- table(x)
  tap <- tapply(y,x,mean,na.rm=T)
  Ylim <- c(0, max(tap)*ymult)
  b <- bargraph.CI(x,y,ylim=Ylim,err.width=0.04,...)

@@

par(mfrow=c(2,2), mar=c(5,5,2,2))
with(pipe_bh, BARplot(pft,alf_stbh, xlab="Plant functional type", 
                       ylab=expression(A[L]/A[W]~breast~height~~(m^2~m^-2)),
                       col=Cols))
with(pipe_bh, BARplot(pft,mlf_stbh, xlab="Plant functional type",
                       ylab=expression(M[F]/A[W]~breast~height~~(kg~m^-2)),
                       col=Cols
                       ))
with(pipe_ba, BARplot(pft,alf_stba, xlab="Plant functional type", 
                          ylab=expression(A[L]/A[W]~base~~(m^2~m^-2)),
                          col=Cols))
b <- with(pipe_ba, BARplot(pft,mlf_stba, xlab="Plant functional type",
                          ylab=expression(M[F]/A[W]~base~~(kg~m^-2)),
                          col=Cols,ymult=1.3
))

@@

@@

par(cex.lab=1.2, mar=c(5,5,2,2))
allombygroupandpftplot("m.to", "m.lf", baad2, legend=T)
allombygroupandpftplot("m.to", "a.lf", baad2)

allombygroupandpftplot("m.so", "m.lf", baad2)
allombygroupandpftplot("m.so", "a.lf", baad2)

allombygroupandpftplot("a.stbh", "m.lf", baad2)
allombygroupandpftplot("a.stba", "m.lf", baad2)

@@

allombygroupandpftplot("m.so", "m.rt", baad2, legend=T)
allombygroupandpftplot("m.lf", "m.rf", baad2, legend=T)

allombygroupandpftplot("c.d", "m.lf", baad2, legend=T)
allombygroupandpftplot("c.d", "d.cr", baad2, legend=T)

@@

@@

@@

>>>>

```{r echo=FALSE}
source("load.R")

@@

pipe_bh <- summarize(group_by(dat, dataset, species), 
           alf_stbh=mean(a.lf / a.stbh, na.rm=TRUE),
           mlf_stbh=mean(m.lf / a.stbh, na.rm=TRUE),
           sla = mean(a.lf / m.lf, na.rm=TRUE),
           pft=first(pft),
           MAP=mean(MAP),
           MAT=mean(MAT),
           growingCondition=first(growingCondition),
           vegetation=first(vegetation))

# Basal pipemodel relationships
datl <- studyWithVars(baad, c("a.stba", "a.lf","m.lf"), returnwhat="list")
dat <- rbind_all(datl)

pipe_ba <- summarize(group_by(dat, dataset, species), 
                     alf_stba=mean(a.lf / a.stba, na.rm=TRUE),
                     mlf_stba=mean(m.lf / a.stba, na.rm=TRUE),
                     pft=first(pft),
                     MAP=mean(MAP),
                     MAT=mean(MAT),
                     growingCondition=first(growingCondition),
                     vegetation=first(vegetation))

dat$pft <- as.factor(dat$pft)
pipe_bh$pft <- as.factor(pipe_bh$pft)
pipe_ba$pft <- as.factor(pipe_ba$pft)

@@

  N <- table(x[!is.na(x)])
  tap <- tapply(y,x,mean,na.rm=T)
  Ylim <- c(0, max(tap)*ymult)
  b <- bargraph.CI(x,y,ylim=Ylim,err.width=0.04,...)

@@

par(mfrow=c(2,2), mar=c(5,5,2,2))
with(pipe_bh, BARplot(pft,alf_stbh, xlab="Plant functional type", 
                       ylab=expression(A[L]/A[W]~breast~height~~(m^2~m^-2)),
                       col=Cols))
with(pipe_bh, BARplot(pft,mlf_stbh, xlab="Plant functional type",
                       ylab=expression(M[F]/A[W]~breast~height~~(kg~m^-2)),
                       col=Cols
                       ))

with(pipe_ba, BARplot(pft,alf_stba, xlab="Plant functional type", 
                          ylab=expression(A[L]/A[W]~base~~(m^2~m^-2)),
                          col=Cols))
with(pipe_ba, BARplot(pft,mlf_stba, xlab="Plant functional type",
                          ylab=expression(M[F]/A[W]~base~~(kg~m^-2)),
                          col=Cols,ymult=1.3
))

par(mfrow=c(1,1))

@@

  legend("bottomright", as.expression(bquote(Slope==.(slopecomp))),
         bty="n", col="grey", cex=0.7)

@@

par(cex.lab=1.2, mar=c(5,5,2,2))
allombygroupandpftplot("m.to", "m.lf", baad2, legend=T, slopecomp=3/4)
allombygroupandpftplot("m.to", "a.lf", baad2, slopecomp=3/4)

allombygroupandpftplot("m.so", "m.lf", baad2, slopecomp=3/4)
allombygroupandpftplot("m.so", "a.lf", baad2, slopecomp=3/4)

allombygroupandpftplot("a.stbh", "m.lf", baad2, slopecomp=1)
allombygroupandpftplot("a.stba", "m.lf", baad2, slopecomp=1)

allombygroupandpftplot("d.bh", "h.t", baad2, slopecomp=1)
allombygroupandpftplot("d.ba", "h.t", baad2, slopecomp=1)

@@

allombygroupandpftplot("c.d", "m.lf", baad2, legend=T, slopecomp=1)
allombygroupandpftplot("c.d", "d.cr", baad2, legend=T, slopecomp=1)

@@

### Root-shoot ratios

Functional balance assumption : does fine root mass scale with leaf mass?
Total root-shoot ratio; would be nice if we find approximate isometry.

```{r echo=FALSE, warning=FALSE, message=FALSE}
allombygroupandpftplot("m.so", "m.rt", baad2, legend=T)
allombygroupandpftplot("m.lf", "m.rf", baad2, legend=T)
```

@@

### Tropical vs. temperate evergreens

```{r echo=FALSE, message=FALSE, warning=FALSE}
ii <- which(P$pft == "EA" & P$vegetation %in% c("TempF","TempRF","TropRF","TropSF"))
l <- l[ii]
P <- P[ii,]

P$TempTrop <- as.factor(ifelse(P$vegetation %in% c("TempF","TempRF"), "Temp", "Trop"))
palette(c("blue","red"))

# log-log scale
plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
     axes=FALSE)
addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
axis(2)
for(i in 1:length(l)){
  points(log10(l[[i]]$h), l[[i]]$wfrel, type='l',
         col=P$TempTrop[i])
}
legend("bottomleft", c("Temperate evergreen angiosperm",
                       "Tropical evergreen angiosperm"), fill=palette())

plot(1, xlim=c(-3,2), ylim=c(-7,5), type='n',
     xlab="Plant height (m)", ylab="Leaf mass (kg)",
     axes=FALSE)
magaxis(side=1:2, unlog='xy',box=TRUE)
for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$mlf), type='l',
                            col=P$TempTrop[i])
legend("topleft", c("Temperate evergreen angiosperm",
                       "Tropical evergreen angiosperm"), fill=palette(),inset=0.03)

plot(1, xlim=c(-3,2), ylim=c(-7,5), type='n',
     xlab="Plant height (m)", ylab="Above-ground woody mass (kg)",
     axes=FALSE)
magaxis(side=1:2, unlog='xy',box=TRUE)
for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$mst), type='l',
                            col=P$TempTrop[i])
legend("topleft", c("Temperate evergreen angiosperm",
                       "Tropical evergreen angiosperm"), fill=palette(),inset=0.03)

@@

### Plots of all data

Something like this for first Figure in manuscript.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- baad

## Fill sapwood area. Let basal=breast height where only data at breast height exists.
# For small plants treat total stem cross section as sapwood
data[["sap"]] <- data[["a.ssba"]]
i <- is.na(data[["sap"]])
data[["sap"]][i] <- data[["a.ssbh"]][i]
i <- is.na(data[["sap"]]) & !is.na(data[["a.stba"]]) & !is.na(data[["h.t"]]) & data[["h.t"]] < 2
data[["sap"]][i] <- data[["a.stba"]][i]

addIsoclines <- function(A= seq(-10,10),b=1,...){
  for(a in A) abline(a,b,lty="dotted",...)
}

add.sma.by.group <-function (data, xvar, yvar, smagroup, colorBygroup, col.table,nmin=10,...){
  if(sum(!is.na(data[[yvar]]*data[[xvar]]))>nmin){
		col=colour.by.category(data[[colorBygroup]][1],col.table)
		fit <- sma(data[[yvar]]~data[[xvar]]*data[[smagroup]], log="xy",quiet=TRUE)
		plot(fit, type='l',col=col,p.lines.transparent=0.1,add=TRUE,...)
		invisible(fit)
	}
}

make.color.table <- function(names){
	col.table <- niceColors(length(names))
	names(col.table) <- names
	col.table
}

# make a plot, fits lines to each within each group, applyiong same colour within that group
plot_with_sma_by_var_within_group <- function(data, xvar,yvar,group="vegetation",location=NULL, isoclines=NULL,...){

	plot(data[[xvar]], data[[yvar]],  type= 'n', log="xy", las=1, yaxt="n", xaxt="n",...)
	axis.log10(1)
	axis.log10(2)

	if(!is.null(isoclines))
		addIsoclines(b=isoclines)

	points(data[[xvar]], data[[yvar]],  type= 'p', col = make.transparent("grey", 0.4), pch=16)

	col.table <- make.color.table(unique(data[[group]]))

	fits <- dlply(data, group, add.sma.by.group, xvar=xvar, yvar=yvar, col.table=col.table,
		smagroup="species",colorBygroup=group)

	get.significant.groups <- function(fit, p=0.1){
		if(is.null(fit)) return(FALSE)
		any(fit$groupsummary$pval < p)
	}

	tmp <- laply(fits, get.significant.groups)
	lab <- sort(names(tmp)[tmp])

	if(!is.null(location))
		legend(location, legend = tolower(lab), bty = "n", pch=16, col=colour.by.category(lab,col.table), cex=0.5)
	fits
}

myfigure <- function(){
# 	par(mfrow = c(1,3))
	fits<-plot_with_sma_by_var_within_group(data,"a.lf", "h.t", "vegetation",
                                          ylim=c(0.005, 120), xlim=c(1E-5, 1E3), 
                                          ylab="Height (m)", 
                                          xlab="",location="topleft", isoclines=0.5, main="a) Architecture")
	legend("bottomright", legend = "Slope = 0.5", bty = "n", cex=1)

	plot_with_sma_by_var_within_group(data,"a.lf", "sap", "vegetation",xlim=c(1E-5, 1E3), ylim=c(1E-8, 1E0), 
                                    ylab=expression(paste("Sapwood area (",m^2,")")), 
                                    xlab=expression(paste("Leaf area (",m^2,")")),location="topleft", 
                                    isoclines=1, "b) Pipe Model")
	legend("bottomright", legend = "Slope = 1", bty = "n", cex=1)

	plot_with_sma_by_var_within_group(data,"a.lf", "m.rt", "vegetation",xlim=c(1E-5, 1E3), ylim=c(1E-8, 400), 
                                    ylab="mass root (kg)", 
                                    xlab="",location="topleft", isoclines=1, main="c) Roots")
	legend("bottomright", legend = "Slope = 1", bty = "n", cex=1)

}

myfigure()
```
